---
title: "DATA607 Project 1"
author: "Gabriel Campos"
date: "`r format(Sys.Date(), '%B %d %Y')`"
output: 
  html_document:
    includes:
      in_header: header.html
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r, echo=FALSE,warning=FALSE}
library(stringr)
library(tidyr)
library(dplyr)
```


Assignment Requirements
=======================




Procedure
=========

## Import Data

Data is imported from the text file path leading to `tournamentinfo.txt` from my [github Project 1 folder](https://github.com/gcampos100/DATA607Spring2021/tree/main/Projects/Project%201). Function `read.delim` from the `utils` package is used. 

```{r}
# Store Github url to variable
txtfile= 
"https://raw.githubusercontent.com/gcampos100/DATA607Spring2021/main/Projects/Project%201/
tournamentinfo.txt"
```

The text file is converted into a data.frame separated by the delimiter `|` and not claiming a `header`, resulting in 11 columns vs. the actual 10 in the file.

```{r}
my_data <- as.data.frame(read.delim(txtfile,header = FALSE,stringsAsFactors = FALSE, sep = "|"))
```

Note[^1]

[^1]:Column 10 was ommitted from head() for formatting:
```{r, echo=FALSE}
head(my_data[,1:9],4)
```

## Cleaning Imported Data

### Step 1: Remove `---` rows and `NULL` values

The data.frame my_data is a 2 dimensional data.frame containing `11 columns` and `195 rows`. Rows composed completely of `---` and the additional all `NULL` column created on import needs to be removed.

```{r}
#Find rows to delete, multiples of 3
toDelete    <- seq(1, length(my_data$V1), 3)
#remove rows
my_data     <- my_data[-toDelete ,]
# remove column with NULL Values
my_data[11] <-NULL
```


Note[^2]

[^2]:Column 9 & 10 was ommitted from head() for formatting:
```{r, echo=FALSE}
head(my_data[,1:8],3)
```

### Step 2: Subset header and body

The header is composed of 2 rows that require merging and separation by column. In order to do so and to avoid affecting the additional data, I subset the first two rows individually and separate the table's body content.

```{r}
# Subset column names and remainder of vectors
# Subset row 1 of my data
my_data_names <- my_data[1,]
# Subset row 2 of my data
my_data_names_2 <- my_data[2,]
# Subset remainder of my data
my_data<-my_data[3:NROW(my_data),]
```

### Step 3: Use header subsets to create column names

Using `str_replace` in combination with `gsub()` and some `regex` expressions, I am able to remove the unnecessary information from the rows then combine them. The end result will be the exact or near exact column names needed for this dataset.

```{r}
# Merge both vectors and seperate by `,`
my_data_names <- paste(my_data_names,my_data_names_2,collapse = ",")
# Replace any character issues
my_data_names <- str_replace(my_data_names, "/",",")
my_data_names <- str_replace(my_data_names, "->"," , ")
my_data_names <- str_replace(my_data_names, "\\(","")
my_data_names <- str_replace(my_data_names, "Post\\)","Rtg Post")
# Remove excessive spacing
my_data_names <-str_replace(gsub("\\s\\s", " ", str_trim(my_data_names)), "B", "b")
# Add comma to attribute `Player Name`
my_data_names <-str_replace(gsub("Player Name", "Player Name,",my_data_names),"B", "b")
```


```{r, echo=FALSE}
my_data_names<-unlist(strsplit(my_data_names, ","))
```

### Step 4 : Use body subset to create columns

The relevant data is split among 2 rows, $\therefore$ I separated the data set into 2 halves in order to rejoin in combined columns. Similar as with the column names, I used a `sequence` to store the `index` of all odd number `indexes` in the `data frame` and created `subsets` of the relevant data.

```{r}
# Subset of top using index
top    <- seq(1, length(my_data$V1), 2)
my_data_top         <- my_data[top ,]
# Subset of bottom using index
bottom     <- seq(2, length(my_data$V1), 2)
my_data_bottom          <- my_data[bottom ,]

```

Note[^3]

[^3]:Columns 1 & 10 was ommitted from head() for formatting:
```{r, echo=FALSE}
paste0("TOP")
head(my_data_top[,2:9],2)
paste0("BOTTOM")
head(my_data_bottom[,2:9],2)
```


### Step 5 : First Data cleanup and consolidation

In order to merge the data cleanly and to allow easy edit of output, I:

* converted each `subset` into a `list`
* Removed data in `my_data_bottom` not needed for the assignment
* Combined the rows into a single column
* Then cleanup or add any characters or symbols that is inhibiting the creating of a table

```{r}
# List conversion
my_data_top <- as.list(t(my_data_top))
my_data_bottom<- as.list(t(my_data_bottom))
```

Note[^4]

[^4]:Regex was used to add `|` character. Although not very intuitive, the pasting of data created issues separating column 1 `Pair Num` and `State`, hence `|` was added to use as a delimeter later on:
```{r}
# Unnecessary Data removal
my_data_bottom<-str_replace_all(gsub("W|B|b|P[1-9]|P[1-9][0-9]", " ",my_data_bottom),"B", "b")
my_data_bottom<-str_replace_all(gsub("O", "| O",my_data_bottom),"B", "b")
my_data_bottom<-str_replace_all(gsub("M", "| M",my_data_bottom),"B", "b")
my_data_bottom<-str_replace_all(gsub("N\\:.", " ",my_data_bottom),"B", "b")
```


```{r}
# clean paste or merger of rows
my_data <- paste(my_data_top,my_data_bottom,collapse = ",")
```


```{r}
# Overall cleanup of data
my_data <-str_replace(gsub("/ R:", ",",my_data),"B", "b")
my_data <-str_replace(gsub("->", ",",my_data),"B", "b")
my_data <-str_replace(gsub("\\s+", " ", str_trim(my_data)), "B", "B")
```

The result of the cleanup is a list of all variables, comma separated, **WITH EXCEPTION** of `Player Name` and `USCF ID`.

```{r, echo=FALSE}
str(my_data)
```

### Step 6: Data Frame recreation

```{r, include=FALSE}
#test.matrix<-matrix(unlist(test.split), ncol = 12, byrow = TRUE)
```

Using the list I recreate the data body data frame with column name data

```{r}
my_data<-strsplit(my_data[1],",")
my_data <- data.frame(matrix(unlist(my_data), ncol=12, byrow=TRUE))
```

Note[^5]

[^5]:Column 10-12 was ommitted from head() for formatting:
```{r, echo=FALSE}
head(my_data[,1:9],2)
```

Specifically column `X2` needs to be separated using `str_split_fixed`
```{r, echo = FALSE}
paste0(my_data$X2[1])
paste0(my_data$X2[2])
```

```{r}
quick.split<-str_split_fixed(my_data$X2, "\\s[0-9]", 2)
my_data<-cbind(my_data[1],quick.split,my_data[3:12])
```

This will to create the 13 columns we require with `my_data_names`

```{r, echo=FALSE}
head(my_data,1)
```

Unfortunately, the `State` column name was not accounted for with `Pair Num` 's aggregate data, hence we have to seperate the first column `X1` using the below command.

Note[^6]

[^6]:This is where the added `|` is used as a delimeter:
```{r}
my_data<-my_data %>%
  separate(X1, c(my_data_names[1], "State"), "\\|")
```

Then the remaining names can be added using the list `my_data_names`

```{r}
colnames(my_data)[colnames(my_data)
            %in% c("1", "2","X3","X4","X5","X6","X7","X8","X9","X10","X11","X12")] <- my_data_names[2:13]
```


```{r, echo=FALSE}
head(my_data,1)
my_data <- apply(my_data,2,toupper)
```


Calculating Average
--------------------

I created a vector Average to store the calculated values that will be put into the data frame later on.

```{r, results=FALSE}
Average<-c(1:64)
```

```{r, echo=FALSE}
dim(my_data)
```


The logic for the below `for loop` is as follows
* Initial loop goes from the first row `1` to the maximum row `nrow`
* Initial column for each row is `7` indexing the `Round 1` column to the maximum column `ncol`.
* Index `i` is retrieved as a numeric value
* `i` is used to index the `row` of opponents data, with `4` indexing `Rtg Pre` value as a numeric
* Value is calculated into `sum` initialized in the first `for loop` as 0, if it passes the conditional `if statement` of not `NA` $\\!is.na$.
* Else statement, tallies number of `NA` values.
* Upon completion of the inner loop, an average for that row is calculated and stored to vector `Average` indexing the same row that is calculated.
* Average calculation = $round(\frac{sum}{column_{\#of}-NA_{\#of}})$

```{r, results=FALSE}
for(row in 1:nrow(my_data)){
    sum = 0
    columns.na=0
  for(col in 8:ncol(my_data)){
    i<-as.numeric(str_extract(my_data[row,col],"[1-9]."))
        if(!is.na(i)){
      print(i)
      sum = sum + as.numeric(my_data[i,5])
        }
      else columns.na=columns.na+1
  }
  #print(columns.na)
  #print(sum)
  #print(round((sum/((ncol(my_data)-7)-columns.na))))
  Average[row]<-round((sum/((ncol(my_data)-7)-columns.na)))
}
```

Once the loop ends and the `Average` vector is filled, I use `cbind()` to combined the data frame and vector by column

```{r}
my_data<-cbind(my_data,Average)
```

```{r, echo=FALSE}
head(my_data[,1:15],3)
```

Convert to CSV
--------------

The requirements for the submission is as follows

      Player’s Name, Player’s State, Total Number of Points, Player’s 
      Pre-Rating, and Average Pre Chess Rating of Opponents

$\therefore$ I subset the data before converting to a `.csv`

```{r}
colnames(my_data)
```

```{r}
#my_data<-as.data.frame(my_data)
Project_1_csv<-as.data.frame(my_data) %>%
  select(" Player Name",State,"Total Pts "," Rtg Pre ","Average")
head(Project_1_csv)
```

After which, I create a .csv file from the data frame using a local github path

```{r}
cloned_git_PATH <- rstudioapi::askForPassword("Enter Path:")
```

```{r}
write.csv(Project_1_csv,
          cloned_git_PATH)
```
