---
title: "DATA607 Week 7 Assignment"
author: "Gabriel Campos"
date: "`r format(Sys.Date(), '%B %d %Y')`"
output: 
  html_document:
    includes:
      in_header: header.html
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, results = "hold")
library(rvest)
library(dplyr)
library(XML)
library(RCurl)
library(janitor)
library(stringr)
```

Overview
========

1. Pick **three** of your favorite books on **one** of your favorite subjects. 
    + **At least one** of the books should have **more than one author**. 
2. For **each book**, include the **title**, **authors**, and **two or three** other **attributes** that you find interesting.
3. Take the information that you’ve selected about these three books, and separately **create three files** which store the book’s information in **HTML** (using an **html table**), **XML**, and **JSON formats** (**e.g.** $“books.html”, “books.xml”, and “books.json”$).
4. To help you better understand the different file structures, I’d prefer that you **create each of these files “by hand”** unless you’re already very comfortable with the file formats.
5. Write `R` code, using **your** `packages` of choice, to **load the information** from each of the **three sources** into separate `R` `data frames`. 
    + **Are the three data frames identical?**
6. Your **deliverable** is the **three source files** and the `R` **code**.
    + If you can, `package` your assignment solution up into an `.Rmd` file and **publish** to $rpubs.com$.       * [This will also require finding a way to make your three text files accessible from the web].

Import Data
===========

[HTML code editor](https://htmlcodeeditor.com/)

```{r, include=FALSE}
url_HTML  <-"https://raw.githubusercontent.com/gcampos100/DATA607Spring2021/main/Week%207/test.html"
urlXML    <-httr::GET("https://raw.githubusercontent.com/gcampos100/DATA607Spring2021/main/Week%207/test.xml")
url_JSON  <-"https://raw.githubusercontent.com/gcampos100/DATA607Spring2021/main/Week%207/test.json"
```

```{r}
import_html <-read_html(url_HTML, header = TRUE)
import_xml <- xmlParse(urlXML)
json_test <-jsonlite::fromJSON(url_JSON)
```


[XMLInternalDocument](https://www.rdocumentation.org/packages/XML/versions/3.99-0.5/topics/XMLInternalDocument-class)

xml [xml_document](https://www.rdocumentation.org/packages/xml2/versions/1.3.2/topics/xml_document-class) class

Convert to Data Frame
=====================

```{r}
df_html<-
  as.data.frame(html_table(import_html)) %>%
    row_to_names(1)

df_xml<-xmlToDataFrame(import_xml)
colnames(df_xml)<- str_to_title(colnames(df_xml))
```

```{r}

```


Cleanup
=======

