---
title: "DATA607 Week 5 Lab"
author: "Gabriel Campos"
date: "`r format(Sys.Date(), '%B %d %Y')`"
output: 
  html_document:
    includes:
      in_header: header.html
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, results = FALSE)
library(tidyverse)
library(nycflights13)
packageDescription("nycflights13")
```

```{r, echo=FALSE}
nycflights13::flights
```

```{r, results='hold'}
(jan<- filter(flights,month == 1, day == 1))
```

Example of approximation
========================

```{r}
sqrt(2)^2 == 2
#vs
near(sqrt(2)^2, 2)
```

AND , OR Operators
==================

```{r, results='hold'}
filter(flights,month == 11 | month == 12)
```

Short-Hand
===========

Note[^1]

[^1]:[from data transformation doc](https://r4ds.had.co.nz/transform.html):
A useful short-hand for this problem is x %in% y. This will select every row where x is one of the values in y. We could use it to rewrite the code above:

```{r, results='hold'}
(nov_dec <- filter(flights, month %in% c(11,12)))
```

Note[^a]

[^a]:[from data transformation doc](https://r4ds.had.co.nz/transform.html):
Sometimes you can simplify complicated subsetting by remembering De Morgan’s law: !(x & y) is the same as !x | !y, and !(x | y) is the same as !x & !y. For example, if you wanted to find flights that weren’t delayed (on arrival or departure) by more than two hours, you could use either of the noted two filters

```{r, results='hold'}
filter(flights, !(arr_delay>120 | dep_delay > 120))
filter(flights, arr_delay <= 120, dep_delay <= 120)
```

Handling NA's
=============

```{r, results='hold'}
(df<- tibble(x =c(1,NA,3)))
filter(df,x>1)
filter(df,is.na(x)|x>1)
```

Exercises
=========

1. Find all flights that:

  + Had an arrival delay of two or more hours
  + Flew to Houston (IAH or HOU)
  + Were operated by United, American, or Delta
  + Departed in summer (July, August, and September)

```{r}
colnames(flights)
```

```{r}
# Regex to check if and combination other than DL exists
str_view(flights$carrier, "^D.", match = TRUE)
str_view(flights$carrier, "^D(?!L)", match = TRUE)
```


```{r}
df_f<-filter(flights, arr_delay > 2, dest == "IAH"| dest =="HOU", carrier == "UA"|carrier == "DL"|carrier == "AA")
df_f[,c("arr_delay","dest","carrier")]
```

## Using `between()` Function

2. Departed between midnight and 6am (inclusive)

```{r}
# Regex to check if and combination other than DL exists
#str_view(flights$dep_time, "^12|^6", match = TRUE)
str_view(flights$dep_time, "^[1-2][3-9][1-9][1-9]|^[2][1-9][1-9][1-9]", match = TRUE)
```

```{r}
filter(flights,between(dep_time, 0000,600))
```

