---
title: "Data Science In Context"
author: "Gabriel Campos"
date: "`r format(Sys.Date(), '%B %d %Y')`"
output: 
  html_document:
    includes:
      in_header: header.html
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, results = "hold")
library(openintro)
library(kableExtra)
```

Overview
========

Data storytelling is a fundamental part of Data Science. The processing and *sense making* of data, is what makes the overload of available today useful and what drives the demand in this field. Although many tools and resources are available for our disposal presently, I would like simply to elaborate on some of the data visualization tools available both internal and external from `R` and the data requirements in creating them. The charts created will be my own from previous work or examples found online.

As usual, plots can be categorized as either exploratory or explanatory, however the focus here is not telling the story of each data set, rather simply a visual presentation of various visualizations, for them.

ggplot
=========

## Summary

ggplot2 is a suite for tools for creating plots. The data used in creating the below ggplot comes from the `tidyverse` package.Details and links regarding `ggplot` can be found below.

(a) `ggplot2` website : [rdocumentation.org/packages/ggplot2/versions/3.3.3](https://www.rdocumentation.org/packages/ggplot2/versions/3.3.3) 
(b) `ggplot` Github: **[github.com/cran/ggplot2](https://github.com/cran/ggplot2)**



### Import Data

```{r, warning=FALSE}
library(ggplot2)
```

### Examples

The data being used comes from [kaggle.com](https://www.kaggle.com/) dataset and .csv file ["Homicide Reports, 1980-2014](https://www.kaggle.com/murderaccountability/homicide-reports?select=database.csv)

```{r, include=FALSE}
file_path <- "D:/Bulk Data/Excel Files/database.csv"
```


```{r}
my_data <- read.csv(file_path)
```

```{r}
my_data %>%
  head() %>%
    kbl() %>%
      kable_styling()

```



