---
title: "DATA607 Week 1 Assignment - Loading Data into a Data Frame"
author: "Gabriel Campos"
date: "2/2/2021"
output: 
  html_document:
    includes:
      in_header: header.html
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r, echo=FALSE,warning=FALSE}
library(kableExtra)
```


Assignment Requirements
=======================
1. Using the 173 majors listed in fivethirtyeight.com’s College Majors dataset [https://fivethirtyeight.com/features/the-economic-guide-to-picking-a-college-major/], provide code that identifies the majors that contain either "DATA" or "STATISTICS"

2. Write code that transforms the data below:

    [1] "bell pepper"  "bilberry"     "blackberry"   "blood orange"

    [5] "blueberry"    "cantaloupe"   "chili pepper" "cloudberry"  

    [9] "elderberry"   "lime"         "lychee"       "mulberry"    

    [13] "olive"        "salal berry"

Into a format like this:

    c("bell pepper", "bilberry", "blackberry", "blood orange", "blueberry", "cantaloupe", "chili pepper", "cloudberry", "elderberry", "lime", "lychee", "mulberry", "olive", "salal berry")

**The two exercises below are taken from R for Data Science, 14.3.5.1 in the on-line version:**

3. Describe, in words, what these expressions will match:

      (.)\1\1
      "(.)(.)\\2\\1"
      (..)\1
      "(.).\\1.\\1"
      "(.)(.)(.).*\\3\\2\\1"

4. Construct regular expressions to match words that:

    Start and end with the same character.
    Contain a repeated pair of letters (e.g. "church" contains "ch" repeated twice.)
    Contain one letter repeated in at least three places (e.g. "eleven" contains three "e"s.)

Overview
========

`The Economic Guide To Picking A College Major` article data is stored on the [fivethirtyeight github](https://github.com/fivethirtyeight/data/tree/master/college-majors). I will be importing my data from the [majors=list.csv](https://github.com/fivethirtyeight/data/blob/master/college-majors/majors-list.csv) and the [grad-students.csv](https://github.com/fivethirtyeight/data/blob/master/college-majors/grad-students.csv) respectively.

```{r}
majors_list<-
  read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/majors-list.csv", header= TRUE, sep=",")

majors_list<-majors_list[order(majors_list$Major),]

grad_students<-
  read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/college-majors/grad-students.csv", header= TRUE, sep=",")

grad_students<-grad_students[order(grad_students$Major),]
```

```{r, echo=FALSE}
paste0("Columns for major_lists are:")
names(majors_list)
paste0("Columns for grad_students are:")
names(grad_students)
```

```{r}
dplyr::filter(majors_list, grepl('data|Data|DATA|statistic|Statistic|STATISTIC', Major))

dplyr::filter(majors_list, grepl('data|Data|DATA|statistic|Statistic|STATISTIC', Major))
```

