---
title: "DATA607 Week 10 Assignment"
author: "Gabriel Campos"
date: "`r format(Sys.Date(), '%B %d %Y')`"
output: 
  html_document:
    includes:
      in_header: header.html
    css: ./lab.css
    highlight: pygments
    theme: cerulean
    toc: true
    toc_float: true
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(eval = TRUE, results = "hold")
library(dplyr)
library(ggplot2)
library(gutenbergr)
library(janeaustenr)
library(reactable)
library(reshape2)
library(stringr)
library(textdata)
library(tidyr)
library(tidytext)
library(tidyverse)
library(wordcloud)
```

Overview
========

**Week 10 Assignment**

Attached Files:

[Week 10 Assignment Rubric.pdf (45.194 KB)](https://bbhosted.cuny.edu/bbcswebdav/pid-54556511-dt-content-rid-425095250_1/xid-425095250_1)\newline

* In *Text Mining with R*, [Chapter 2 looks at Sentiment Analysis](https://www.tidytextmining.com/sentiment.html).  In this assignment, you should start by     getting the primary example code from chapter 2 working in an R Markdown document.  You should provide a
  citation to this base code.  You’re then asked to extend the code in two ways:
*	Work with a different corpus of your choosing, and
*	Incorporate at least one additional sentiment lexicon (possibly from another R package that you’ve found
  through research).

$As\ usual,\ please\ submit\ links\ to\ both\ an\ .Rmd\ file\ posted\ in\ your\ GitHub\ repository\ and\ to\ your\ code\ on\ rpubs.com.$\newline $\ You\ make\ work\ on\ a\ small\ team\ on\ this\ assignment.$

# Chapter 2 example {.tabset .tabset-pills} 

## 2.1

**The `sentiments` datasets **


```{r, results='hold'}
get_sentiments("afinn")
```

```{r, results='hold'}
get_sentiments("bing")
```

```{r, results='hold'}
get_sentiments("nrc")
```

## 2.2

**Sentiment analysis with inner join**

```{r}
tidy_books <-
  austen_books() %>%
    group_by(book) %>%
      mutate(linenumber = row_number(),
              chapter = 
               cumsum(str_detect(text,
                      regex("^chapter [\\divxlc]",
                        ignore_case = TRUE)))) %>%
                                          ungroup() %>%
                                            unnest_tokens(word, text)
```

```{r, results='hold'}
nrc_joy <- 
  get_sentiments("nrc") %>% 
    filter(sentiment == "joy")

tidy_books %>%
  filter(book == "Emma") %>%
    inner_join(nrc_joy) %>%
      count(word, sort = TRUE)
```

```{r}
jane_austen_sentiment <- 
  tidy_books %>%
    inner_join(get_sentiments("bing")) %>%
      count(book,index = linenumber %/% 80,sentiment) %>%
        pivot_wider(names_from = sentiment,values_from = n,
                                          values_fill = 0) %>% 
          mutate(sentiment = positive - negative)
```

```{r}
ggplot(jane_austen_sentiment, aes(index, sentiment, fill = book)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~book, ncol = 2, scales = "free_x")
```

## 2.3

**Comparing the three sentiment dictionaries**

```{r, results='hold'}
pride_prejudice <- 
  tidy_books %>% 
    filter(book == "Pride & Prejudice")
```

```{r}
afinn <- pride_prejudice %>% 
  inner_join(get_sentiments("afinn")) %>% 
      group_by(index = linenumber %/% 80) %>% 
          summarise(sentiment = sum(value)) %>% 
            mutate(method = "AFINN")
```

```{r}
bing_and_nrc <- 
  bind_rows(pride_prejudice %>% 
    inner_join(get_sentiments("bing")) %>%
      mutate(method = "Bing et al."),
              pride_prejudice %>% 
                inner_join(get_sentiments("nrc") %>% 
                  filter(sentiment %in% c("positive", 
                                           "negative"))
                                                  ) %>%
                                mutate(method = "NRC")) %>%
        count(method, index = linenumber %/% 80, sentiment) %>%
                                pivot_wider(names_from = sentiment,
                                            values_from = n,
                                            values_fill = 0) %>% 
                                              mutate(sentiment = positive - negative)
```


```{r}
bind_rows(afinn, 
          bing_and_nrc) %>%
            ggplot(aes(index, sentiment, fill = method)) +
            geom_col(show.legend = FALSE) +
            facet_wrap(~method, ncol = 1, scales = "free_y")
```

```{r}
get_sentiments("nrc") %>% 
  filter(sentiment %in% c("positive", "negative")) %>% 
    count(sentiment)
```

```{r}
get_sentiments("bing") %>% 
  count(sentiment)
```

## 2.4

**Most common positive and negative words**

```{r}
bing_word_counts <- tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
    count(word, sentiment, sort = TRUE) %>%
      ungroup()
```

```{r}
bing_word_counts %>%
  group_by(sentiment) %>%
    slice_max(n, n = 10) %>% 
      ungroup() %>%
        mutate(word = reorder(word, n)) %>%
          ggplot(aes(n, word, fill = sentiment)) +
          geom_col(show.legend = FALSE) +
          facet_wrap(~sentiment, scales = "free_y") +
          labs(x = "Contribution to sentiment",
          y = NULL)
```

```{r}
custom_stop_words <- 
  bind_rows(tibble(word = c("miss"),  
            lexicon = c("custom")), 
            stop_words)
```

## 2.5

**Wordclouds**

```{r}
tidy_books %>%
  anti_join(stop_words) %>%
    count(word) %>%
      with(wordcloud(word,
                     n,
                     max.words = 100))
```

```{r}
tidy_books %>%
  inner_join(get_sentiments("bing")) %>%
    count(word, sentiment, sort = TRUE) %>%
      acast(word ~ sentiment, value.var = "n", fill = 0) %>%
        comparison.cloud(colors = c("gray20"
                                    , "gray80"),
                          max.words = 100)
```

## 2.6

**Looking at units beyond just words**

```{r}
p_and_p_sentences <- 
  tibble(text = prideprejudice) %>% 
  unnest_tokens(sentence,
                text,
                token = "sentences")
```

```{r}
p_and_p_sentences$sentence[2]
```

```{r}
austen_chapters <- 
  austen_books() %>%
    group_by(book) %>%
      unnest_tokens(chapter,
                    text,
                    token = "regex", 
                    pattern = "Chapter|CHAPTER [\\dIVXLC]") %>%
                      ungroup()

austen_chapters %>% 
  group_by(book) %>% 
    summarise(chapters = n())
```

```{r}
bingnegative <-
  get_sentiments("bing") %>% 
  filter(sentiment == "negative")

wordcounts <- tidy_books %>%
  group_by(book, chapter) %>%
    summarize(words = n())

tidy_books %>%
  semi_join(bingnegative) %>%
    group_by(book, chapter) %>%
      summarize(negativewords = n()) %>%
      left_join(wordcounts, by = c("book", "chapter")) %>%
        mutate(ratio = negativewords/words) %>%
          filter(chapter != 0) %>%
            slice_max(ratio, n = 1) %>% 
              ungroup()
```

# Different Corpus {.tabset .tabset-pills}

## Corpus selection 

In order to conduct a sentiment analysis I will select have to select a novel from the [^`gutenbergr`]: This is a random test. package. Its fields can be shown below. Upon selecting a corpus I will import it according to the id #.

```{r, echo=FALSE}
names(gutenberg_metadata)%>%
  kableExtra::kbl(col.names = c("gutenberg fields"))%>%
  kableExtra::kable_styling()
```

```{r}

gutenberg_metadata %>%
            select(gutenberg_id, title,author, gutenberg_bookshelf) %>%
              filter(gutenberg_id>0)%>%
               rename(id = gutenberg_id, subject = gutenberg_bookshelf)%>%
          reactable(
            bordered = TRUE,
            striped = TRUE,
            highlight = TRUE,
            filterable = TRUE,
            showPageSizeOptions = TRUE,
            showPagination = TRUE,
            columns = list(
                    id      = colDef(minWidth = 70),
                    title   = colDef(minWidth = 400),
                    author  = colDef(minWidth = 100),
                    subject = colDef(minWidth = 200)
                    ),
          fullWidth = TRUE,
          defaultPageSize = 5)

```

Ultimately I chose to use the author of `The Legend of Sleepy Hollow` or id # `41`, whose name is `Irving, Washington`.


```{r, message= FALSE, warning= FALSE}
(irv_txt <- gutenberg_download(c(41, 877, 1371, 1850, 2048, 3293, 7002, 7948, 7993, 7994, 8519, 8571, 13042, 13514, 13515, 14228, 19293, 20656, 21195, 32987, 36652, 38192, 49258, 49259, 49872, 49947, 50352)))
```

## Tidy, Bing & NRC {.tabset }

### Sentiment Analysis

**Tidy**

I use `stop_words` in the `anti_join()` function from the `dplyr` package to remove stop words from my download, and tidy my data so that its `gutenberg_id`, `linenumber` and `word` is stored. I then get a count of each word and sort accordingly.


```{r, message = FALSE}
tdy_irv_txt <-
  irv_txt %>%
    mutate(linenumber = row_number()) %>%
              unnest_tokens(word, text) %>%
                anti_join(stop_words)

irv_cnt<-
  tdy_irv_txt %>%
  count(word, sort = TRUE)
```

**Bing**

Using bing, I'm concerned with the sentiment based on the word. Following the example from 2.2 exactly provides the below results

```{r, message = FALSE}
(tdy_irv_bng <- 
  tdy_irv_txt %>%
    inner_join(get_sentiments("bing")) %>%
      count(word,
            index = linenumber %/% 80,
            sentiment) %>%
        pivot_wider(names_from = sentiment,
                    values_from = n,
                    values_fill = 0) %>% 
          mutate(sentiment = positive - negative))
```

I found a full count more useful however

```{r, message=FALSE}
(tdy_irv_bng_cnts <-
  tdy_irv_txt %>%
    inner_join(get_sentiments("bing")) %>%
      count(word, sentiment, sort = TRUE) %>%
        ungroup())
```

**NRC**

The above mentioned `pipe` with `innerjoin` command can also provide a very functional way to utilize `NRC` sentiment.

```{r, message= FALSE}
(nrc_cnts <- 
  tdy_irv_txt %>%
    inner_join(get_sentiments("nrc")) %>%
      count(word, sentiment, sort = TRUE) %>%
        ungroup())
```

if we use the `nrc_joy` filtered sentiment from 2.2 on `gutenberg_id` number `41` for $The\ Legend\ of\ Sleepy\ Hollow$ the results are as follows

```{r}
tdy_irv_txt %>%
  filter(gutenberg_id == 41) %>%
  inner_join(nrc_joy) %>%
  count(word, sort = TRUE)
```

### Comparing 3 Sentiment dictionaries

Here we test the presence of afinnn, bing and nrc in Washington Irvin's work. the afinn and bing_and_nrc variables are overwritten with our new results.

```{r,message= FALSE}
afinn <- 
  tdy_irv_txt %>% 
    inner_join(get_sentiments("afinn")) %>% 
      group_by(index = linenumber %/% 5) %>% 
        summarise(sentiment = sum(value)) %>% 
          mutate(method = "AFINN")
```

```{r,message= FALSE}

bing_and_nrc <- 
  bind_rows(
    tdy_irv_txt %>% 
      inner_join(get_sentiments("bing")) %>%
        mutate(method = "Bing et al."),
          tdy_irv_txt %>% 
            inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
      count(method,
            index = linenumber %/% 5,
            sentiment) %>%
          pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
                mutate(sentiment = positive - negative)
```

```{r, message=FALSE}
bind_rows(afinn, 
          bing_and_nrc) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")
```

From here `gutenberg_id` number `41` for $The\ Legend\ of\ Sleepy\ Hollow$ is again used.

```{r, message=FALSE}
(Sleepy_Hollow<-
  tdy_irv_txt %>%
    filter(gutenberg_id == 41))

```

```{r, message=FALSE}
afinn_sh <- Sleepy_Hollow %>% 
  inner_join(get_sentiments("afinn")) %>% 
  group_by(index = linenumber %/% 5) %>% 
  summarise(sentiment = sum(value)) %>% 
  mutate(method = "AFINN")
```

```{r, message=FALSE}
bing_and_nrc_sh <- bind_rows(
  Sleepy_Hollow %>% 
    inner_join(get_sentiments("bing")) %>%
    mutate(method = "Bing et al."),
  Sleepy_Hollow %>% 
    inner_join(get_sentiments("nrc") %>% 
                 filter(sentiment %in% c("positive", 
                                         "negative"))
    ) %>%
    mutate(method = "NRC")) %>%
  count(method, index = linenumber %/% 5, sentiment) %>%
  pivot_wider(names_from = sentiment,
              values_from = n,
              values_fill = 0) %>% 
  mutate(sentiment = positive - negative)
```

```{r, message=FALSE}
bind_rows(afinn_sh, 
          bing_and_nrc_sh) %>%
  ggplot(aes(index, sentiment, fill = method)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~method, ncol = 1, scales = "free_y")
```

### Most common $+$ & $-$ words

Again over writing variable `bing_word_counts`

```{r, message=FALSE}
(bing_word_counts <- 
  tdy_irv_txt %>%
    inner_join(get_sentiments("bing")) %>%
      count(word, sentiment, sort = TRUE) %>%
        ungroup())
```

```{r, message=FALSE, warning=FALSE}
bing_word_counts %>%
  group_by(sentiment) %>%
  slice_max(n, n = 10) %>% 
  ungroup() %>%
  mutate(word = reorder(word, n)) %>%
  ggplot(aes(n, word, fill = sentiment)) +
  geom_col(show.legend = FALSE) +
  facet_wrap(~sentiment, scales = "free_y") +
  labs(x = "Contribution to sentiment",
       y = NULL)
```

### WordClouds

```{r, message=FALSE, warning=FALSE}
set.seed(528)
tdy_irv_txt %>%
  anti_join(stop_words) %>%
  count(word) %>%
  with(wordcloud(word, n, max.words = 100,
          rot.per=0.35,
          colors=brewer.pal(7, "Accent")))
```


```{r, message=FALSE, warning=FALSE}
tdy_irv_txt %>%
  inner_join(get_sentiments("bing")) %>%
  count(word, sentiment, sort = TRUE) %>%
  acast(word ~ sentiment, value.var = "n", fill = 0) %>%
  comparison.cloud(colors = c("red", "blue"),
                   max.words = 100)
```

# Different Lexicon

For this portion I decided to work with the `Syuzhet Package` and follow the material provided in the [syuzhet-vignette](https://cran.r-project.org/web/packages/syuzhet/vignettes/syuzhet-vignette.html), which involves:

* Retrieving the text format of a text in the form of a string. Raw text source is available on [gutenberg.org](https://www.gutenberg.org/ebooks/)
* Methods available to retrieve the text are:
    + `get_text_as_string()`
    + `get_tokens()`
    + `get_sentences()`
* Use of `get_sentiment()` on the collected sentence or tokens
* Various methods can be used with `get_sentiment()` e.g.:
    + `get_sentiment(token, method="syuzhet")`
    + `get_sentiment(token, method = "bing")`
    + `get_sentiment(token, method = "afinn")`
    + `get_sentiment(token, method = "nrc", lang = "english")`
    + `get_sentiment(sentence_vector)`

For it's usage I will again use $The\ Legend\ of\ Sleepy\ Hollow$ located [here](https://www.gutenberg.org/ebooks/41)

```{r,message = FALSE, warning = FALSE}
# import Library
library(syuzhet)
# Retrieve text
sh_book <- get_text_as_string("https://www.gutenberg.org/files/41/41.txt")
# create sentence vector
str(sh_s_v<- get_sentences(sh_book))
# Create Tokens
book_v <- get_tokens(sh_book, pattern = "\\W")
# create vectors
syuzhet_vector  <- get_sentiment(book_v, method="syuzhet")
bing_vector     <- get_sentiment(book_v, method = "bing")
afinn_vector    <- get_sentiment(book_v, method = "afinn")
nrc_vector      <- get_sentiment(book_v, method = "nrc", lang = "english")
s_v_sentiment   <- get_sentiment(sh_s_v)

```

The heads of which in is shown in the 4x6 matrix below.

```{r,message = FALSE, warning = FALSE}
rbind(
  sign(head(syuzhet_vector)),
  sign(head(bing_vector)),
  sign(head(afinn_vector)),
  sign(head(nrc_vector))
)
```

specifically the sum, mean and summary of the syuzhet_vector is as follows

```{r,warning=FALSE, message=FALSE}
sum(syuzhet_vector)
mean(syuzhet_vector)
summary(syuzhet_vector)
```

My goal was to use the methods noted in the vignette and see how it would change my understanding of the book `Sleep Hollow` if it did at all. `plot()` function was used with both the `syuzhet_vector` and `s_v_sentiment` my sentence vectors as a measure of the negative and positive sentiments throughout the story. X being here a measurement of time, Y varying according to time withing the story.

I unfortunately found this plot the least useful, since as the vignette notes, $'for\ visualization\ it\ is\ generally\ preferable\ to\ remove\ the\ noise\ and\ reveal\ the\ simple\ shape\ of\ the\ trajectory'$

```{r,warning=FALSE, message=FALSE}
plot(
  syuzhet_vector, 
  type="h", 
  main="syuzhet vector Plot Trajectory", 
  xlab = "Narrative Time", 
  ylab= "Emotional Valence"
  )

plot(
  s_v_sentiment, 
  type="l", 
  main="Sentence Vector Plot Trajectory", 
  xlab = "Narrative Time", 
  ylab= "Emotional Valence"
  )
```

I feel the percentage based plot was most useful, indicat a steady rise in negativity in what I beleive to be the climax of the book and a huge shift to a positive valence, in the conclusion.

```{r,warning=FALSE, message=FALSE}
percent_vals <- get_percentage_values(syuzhet_vector, bins = 10)
plot(
  percent_vals, 
  type="l", 
  main="Sleepy Hollow Using Percentage-Based Means", 
  xlab = "Narrative Time", 
  ylab= "Emotional Valence", 
  col="red"
  )
```

Using `Percentage-Base Means` echos the plot above, with more fluctuation which is more inline with how this story is particularly told in my perspective.

```{r,warning=FALSE, message=FALSE}
percent_vals <- get_percentage_values(syuzhet_vector, bins = 20)
plot(
  percent_vals, 
  type="l", 
  main="Sleepy Hollow Using Percentage-Based Means", 
  xlab = "Narrative Time", 
  ylab= "Emotional Valence", 
  col="red"
  )
```

Visually utilization of ft_values to smooth out these representations limits the chaotic perception given from percent_vals.

```{r,warning=FALSE, message=FALSE}
ft_values <- get_transformed_values(
      syuzhet_vector, 
      low_pass_size = 3, 
      x_reverse_len = 100,
      padding_factor = 2,
      scale_vals = TRUE,
      scale_range = FALSE
      )
```

```{r}
plot(
  ft_values, 
  type ="l", 
  main ="Joyce's Portrait using Transformed Values", 
  xlab = "Narrative Time", 
  ylab = "Emotional Valence", 
  col = "red"
  )
```

However the practicality of `simple_plot()` would be my preference in most scenarios.

```{r}
simple_plot(s_v_sentiment)
```

[^a-random-footnote-label]: This is a random test.